\section{The Halting Problem}

\subsection{(Turing Machine) Computable Functions}

\begin{definition}[Compute]\mbox{}\par
Let Language
\(L(M) = \{x \in \Sigma^* | m \text{ accepts } x\}\),
\(\text{Output}(M,x) :=\) the content of the tape of \(M\) when it halt.

We say \(M\) \emph{computes} \(f: \Sigma^* \rightarrow \Sigma^*\) if:
\begin{itemize}
    \item
  \(M\) halts on every \(x\);
\item
  \(\text{Output}(M,x) = f(x)\)
\end{itemize}
\textbf{Computable}: \(f\) is computable if there is a Turing Machine
computing \(f\).
\end{definition}

\begin{theorem}[A Universal Turing Machine exists]
Machine \(U\) over \(\Sigma = \{0,1\}\) such that on input \(x\):
\begin{itemize}
\def\labelenumi{\arabic{enumi}.}
\item
  \(U\) rejects \(x\) if \(x\) is not of the form
  \(\langle M,y\rangle\), \(y \in \{0,1\}^*\)
\item
  else

  \begin{itemize}
  \item
    \(U\) accepts \(x\) \(\Leftrightarrow\) \(M\) accepts \(x\)
  \item
    \(U\) rejects \(x\) \(\Leftrightarrow\) \(M\) rejects \(x\)
  \item
    if \(M\) on \(x\) halts then \(U\) halts with \(\text{Output}(M,x)\)
    on its last tape.
  \end{itemize}
\end{itemize}
\end{theorem}

\begin{definition}[Decidable]
\(L \subseteq \Sigma^*\) is
\emph{decidable} if exists a Turing Machine \(M\) s.t.:

\begin{itemize}
\item
  \(L = L(M)\)
\item
  \(M\) halts on every input
\end{itemize}
\end{definition}

\begin{observation}


Computable examples

\begin{itemize}
\item
  if \(f: \{0,1\}^* \rightarrow \{0,1\}^*\) is primitive recursive then
  \(f\) is computable.
\item
  Ackerman Function is computable.
\item
  If a C++ program can compute \(f\) then \(f\) is computable.
\end{itemize}

Not computable examples

\begin{itemize}
\item
  How many Turing Machines are there? (\(\mathbb{N} \sim \aleph_0\))
\item
  How many languages \(L \subseteq \Sigma^*\) are there?
  (\(2^{\mathbb{N}} \sim \mathbb{R}\)).
\end{itemize}
\end{observation}

\subsection{Kolmogorov Complexity}

\begin{definition}[Kolmogorov complexity]
Kolmogorov complexity measures how random a string is.

Can be formalized as the length of shortest description
\(\langle M\rangle x\), where \(M\) is a Turing Machine description, and
\(x\) is the compressed words. (Like the self-extracting compressed
archive), as

$$KC(w) = \arg\min( \langle M \rangle x| U(M,x) = w)$$

\end{definition}

Example:

\begin{itemize}
\item
  \(\texttt{010101010101} \rightarrow (\texttt{01})^5\)
\item
  \(3.1415... \rightarrow \sqrt{6 \sum_n{\frac{1}{n^2}}} (= \pi)\)
\end{itemize}

\begin{theorem}
Kolmogorov Complexity \(KC(w)\) is not a computable
function.
\end{theorem}

\begin{proof}
\begin{observation}
$KC(x)$ is never much larger than $x$. Otherwise a do-nothing Turing machine and the original can do the job.
\end{observation}

\begin{lemma}
$\forall n, \exists x \in \{0,1\}^n$ with $KC(x) \ge n$.
\end{lemma}

\begin{proof}
At most $2^k$ strings $k$ can be represented with $KC(y)=k$. Then $KC(y)< n$ can represent $2^{n}$.
\end{proof}

If $KC(w)$ is computable, define a function $GKC(n)$ that traverses $\{0,1\}^n$ and return one that $KC(x) \ge n$. A Turing machine computing this will have $|GKC| + \log n$ bits. And finally $|GKC| + \log n < n$. Which contradicts to the definition of $KC$.

\end{proof}

\subsection{The Halting Problem}

Turing Machine M, as encoded $\langle M \rangle$:

$$\text{\textsc{Halt}}:=\{x | M \text{ halts on }x\}$$

\begin{theorem}
Halt is undecidable.
\end{theorem}

\begin{proof}[Proof with diagonalization]
There is no bijection
\(f: \mathbb{N} \rightarrow \{0,1\}^\infty\).

The Turing machine is countable: a surjection that $\mathbb{N}
\rightarrow \{0,1\}^*$

Assume Halt is decidable, then there is a Turing machine $T$ compute the
halt (a diagonal table with $x$ be all Turing machines numbered, $y$ be all Turing machines, and item represents whether it holds)

diagonal negation: Turing machine $\lnot D(table)$:

if x $\not\in \Sigma$ : reject.

if $M$ fails to hold on $x$: accept.

if $M$ holds on $x$: run infinite loop.

Whether $M$ halts on $x$? contradiction.
\end{proof}



\subsection{Post Correspondence
Problem}

\begin{definition}[Post Correspondence Problem]
The definition of PCP:

\begin{itemize}
\item
  A finite set of tile $T$,  $\tile{a}{b} \in T$ with $a,b \in \Sigma^*$.
\item
  For a sequence of tiles $s=s_1s_2\dots s_n$, let
  $top(s)=a_1a_2\dots a_n$,
  $bottom(s)=b_1b_2\dots b_n$
\end{itemize}

Is there a sequence of tiles $S$ with $top(s)=bottom(s)$?
\end{definition}

\begin{theorem}
PCP is undecidable.
\end{theorem}

\begin{proof}
\begin{definition}[MPCP]
Modified PCP, where particular $t^* \in T$, where sequence must start with $t^*$.
\end{definition}

\begin{lemma}
If we can decide PCP, we can decide MPCP. We can turn a MPCP problem into PCP problem.
\end{lemma}

\begin{proof}
We may need to modify MPCP to make sure the first tile should always be the first tile.

The new tile $t^* = \tile{1000}{100}$ becomes $\tile{*1*0*0*0}{*1*0*0*}$. 

Other tiles $\tile{1000}{100}$ becomes $\tile{*1*0*0*0}{1*0*0*}$.

Add a new ending tile $\tile{*\$}{\$}$. Tile $t^*$ needs to be copied as other tiles once.
\end{proof}

\begin{theorem}[Reduction from halting problem to MPCP]
MPCP is undicidable. In particular, given a Turing machine we can translate it into a MPCP over the alphabet $\Gamma \cup Q \cup \{\#\}$ s.t.

$$M \text{ holds on } x \Leftrightarrow MPCP \text{ has a solution}$$
\end{theorem}

\begin{proof}
We construct tiles with configurations.

Start configuration tile: $$\tile{\#\# q_{start}abc... \#}{\#\#}$$

Current sequence tiles patch looks like $$\tile{\#\# C_0 \# \dots \# C_k\# \mathbf{u} a\mathbf{q}bc \mathbf{v} \#}{\#\# C_0 \# \dots \# C_k\#}$$

We define copy tiles which copies all characters $\tile{a}{a}$ for each $a \in \Gamma$.

Rule tiles:
\begin{itemize}
    \item for $(q,m) \rightarrow (q', n, \text{STAY})$:
    
    $$\tile{q'n}{qm}$$
    
    \item for $(q,m) \rightarrow (q', n, \text{RIGHT})$:
    
    $$\tile{nq'}{qm}$$
    \item for $(q,m) \rightarrow (q', n, \text{LEFT})$:
    
    $$\tile{q'An}{Aqm} \text{ for all } A \in \Gamma$$
\end{itemize}

We need ending tiles to collect from the $q_{accept}$ state.

$\tile{q_{accept}}{q_{accept}b}$ for each $b \in \Gamma$, and $\tile{q_{accept}\#}{a q_{accept}\#}$ for each $a \in \Gamma$.

\end{proof}

\end{proof}